Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-07-15 sports_analytics.Rmd
minor edits

In basketball, analytics has driven a shift in the conventional wisdom about shot selection. 
Most teams are shooting more three-pointers, settling for fewer long two-point shots, deploying more versatile defenders, and relying less on the strategy of pounding the ball into the paint in an attempt to get a high-percentage shot. [CITATION NEEDED?]
In American football, teams are going for it on fourth down far more often than in the past, a direct result of statistical analysis showing that most teams were previously overly conservative. [LOPEZ?]
And of course in baseball, teams are using defensive shifts to maximize the probability of recording an out, encouraging hitters to improve their launch angles, and optimizing pitcher repertoires to minimize contact.

@@

A demonstration of the EPV model presented in @cervone2016multiresolution's paper is available on the first author's GitHub at <https://github.com/dcervone/EPVDemo>. As an illustration, using the provided tracking data, Figure \ref{fig:epv-demo} shows a snapshot of a possession during the 2013-14 NBA regular season matchup between the Miami Heat and the Brooklyn Nets on November 1, 2013. Miami is the team on offense in this play, and the possession outcome is a 26-foot three-point miss by Mario Chalmers of the Heat. The plot consists of two elements: 1) (bottom) the player locations on the court at a particular moment in this possession--when the ball just left Chalmers' hands, and 2) (top) a line graph showing how the EPV has continuously changed throughout the play until the three-point attempt. For this entire possession, the estimated EPV for the Miami Heat peaks at this single time point at $1.276$.

>>>>

In basketball, analytics has driven a shift in the conventional wisdom about shot selection. 
Most teams are shooting more three-pointers, settling for fewer long two-point shots, deploying more versatile defenders, and relying less on the strategy of pounding the ball into the paint in an attempt to get a high-percentage shot [@Schuhmann2021nba3point].
In American football, teams are going for it on fourth down far more often than in the past, a direct result of statistical analysis showing that most teams were previously overly conservative. [LOPEZ?]
And of course in baseball, teams are using defensive shifts to maximize the probability of recording an out, encouraging hitters to improve their launch angles, and optimizing pitcher repertoires to minimize contact.

@@

A demonstration of the EPV model presented in @cervone2016multiresolution's paper is available on the first author's GitHub at <https://github.com/dcervone/EPVDemo>. As an illustration, using the provided tracking data, Figure \ref{fig:epv-demo} displays a snapshot of a possession during the 2013-14 NBA regular season matchup between the Miami Heat and the Brooklyn Nets on November 1, 2013. Miami is the team on offense in this play, and the possession outcome is a 26-foot three-point miss by Mario Chalmers of the Heat. The plot consists of two elements: 1) (bottom) the player locations on the court at a particular moment in this possession--when the ball just left Chalmers' hands, and 2) (top) a line graph showing how the EPV is continuously changing throughout the play until the three-point attempt. For this entire possession, the estimated EPV for the Miami Heat reaches its peak at $1.276$ at this specific time point.
